Portable Multi- and Many-Core Performance for Finite Difference Codes; Application to the Free-Surface Component of NEMO.

### 2017-06-20 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolite-paper: add text describing short loop and kernel fusin

When using the OpenACC \texttt{kernels} directive the compiler creates
GPU kernels by parallelising only the outer loop of each loop nest.
For the majority of loop nests in NEMOLite2D all of the iterations are
independent. It is therefore beneficial to expose the maximum amount
of parallelism to the GPU and we do this by instructing the compiler
to parallelise both levels (i.e. \texttt{ji} and \texttt{jj}) of a
loop nest. This is achieved using the OpenACC \texttt{loop collapse(2)}
directive.

@@

As well as accounting for the bulk of the execution time, the Momentum
kernels ($u$ and $v$) are also relatively large.

@@

>>>>

When using the OpenACC \texttt{kernels} directive the compiler creates
GPU kernels by parallelising only the outer loop of each loop nest.
For the majority of loop nests in NEMOLite2D all of the iterations are
independent and the loop bodies consist of just a handful of
executable statements. For small kernels the loop start-up cost can be
significant and therefore it is therefore beneficial to generate as
many threads as will fit on the GPU and then re-use them as
required. For this approach to be efficient we must expose the maximum
amount of parallelism to the GPU and we do this by instructing the
compiler to parallelise both levels (i.e. \texttt{ji} and \texttt{jj})
of a loop nest. This is achieved using the OpenACC \texttt{loop
  collapse(2)} directive for all of the smaller kernels in NEMOLite2D.

@@

In contrast to all of the other NEMOLite2D kernels, the Momentum
kernels ($u$ and $v$) are relatively large in terms of the number of
executable statements they contain. In turn, this means that the
corresponding kernels will require more state and thus more registers
on the GPU. Therefore, rather than generating as many threads as will
fit on the GPU, it is more efficient only to generate as many as
required by the problem. Once slots become free on the GPU new threads
will launch. This minimises the overhead associated with thread
creation and eases register pressure.  The compiler can be persuaded
to adopt the above strategy through the use of the (PGI-specific)
\texttt{loop shortloop} directive which tells it that there is not
likely to be much thread re-use in the following loop. This directive
was applied to both the \texttt{ji} and \texttt{jj} loops in both of
the Momentum kernels.

@@

Both of the Momentum kernels read from 20 double-precision arrays and
thus require considerable memory bandwidth.  However, the majority of
these arrays are used by both the $u$ and $v$ kernels. It is therefore
possible to reduce the memory-bandwidth requirements by fusing the two
kernels together. In practice, this means fusing the two,
doubly-nested loops so that we have a single loop nest containing both
the $u$ and $v$ kernel bodies/calls.
